### 2025-09-25 paper/paper_draft.qmd
Remove use of "category"
@eco-ben , you seem to prefer "cluster" over "category" so to avoid any confusion I have tried to standardize to "cover cluster" when talking about "High/Med/Low" time series clusters.

---
title: Think deep for reef resilience
author:
  - name: Benjamin Grier
    email: grierben777@gmail.com
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          state: QLD
    attributes:
        corresponding: true
  - name: Takuya Iwanaga
    email: t.iwanaga@aims.gov.au
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          city: Cape Cleveland
          state: QLD
    attributes:
        corresponding: true
abstract: |
    Coral reefs globally face unprecedented threats from climate change, with the Great Barrier
    Reef (GBR) experiencing cumulative stressors and increasingly severe declines in coral
    cover from thermal stress events.
    Understanding drivers behind reef resilience to climate impacts is critical for conservation
    planning and intervention strategies. A scenario-based population modelling approach was
    adopted with larval connectivity dynamics, thermal adaptation, and environmental factors
    to assess coral reef resilience across GBR locations under future climate projections.
    Using five Global Climate Models (GCMs) spanning five Equilibrium Climate Sensitivities,
    we analysed reef coral cover time series and the ability of reefs to maintain positive carbonate
    production budgets under future conditions. Contrary to expectations, larval connectivity
    patterns did not correlate with increases in maintenance of positive carbonate budgets.
    Instead, reef depth emerged as the primary predictor of cover-time series resilience and
    maintenance of positive carbonate budgets, with deeper reefs benefiting from thermal
    refugia that reduce bleaching impacts. Interaction between reef depth and incoming
    connectivity was found to influence reef temporal behaviour at 15m depth. Our findings suggest that
    depth provides critical protection against thermal stress, with high incoming connectivity enhancing the
    performance of deep reefs, contrasting current intervention
    practices that primarily target easily accessible sites in the 1-10m depth range. These results have
    important implications for reef management, indicating that depth refugia should be
    prioritized for further consideration in conservation planning alongside connectivity
    considerations to maximize coral survival under continuing climate change.
keywords:
  - Coral reef
  - Climate change
  - Ecosystem resilience
  - Reef depth
  - Reef connectivity
date: last-modified
format:
  elsevier-docx:
    journal:
      name: Journal Name
      formatting: preprint
      cite-style: authoryear
bibliography: references.bib
engine: julia
crossref:
  custom:
  - kind: float
    reference-prefix: SFigure
    key: sfig
  - kind: float
    reference-prefix: STable
    key: stbl
editor:
  markdown:
    wrap: 92
    canonical: true
---

# Introduction

Climate change is impacting species survival, persistence and ecosystem functions globally.
This is particularly true for coral reefs, which are vulnerable to increasing magnitude and
variation in environmental stressors such as ocean temperature, acidification and nutrient
loading [@foster2021; @gattuso2014]. Mass bleaching and mortality events driven by climate
change can cause reef-building coral population declines [@hughes2017; @hughes2018;
@bozec2025], possibly leading to long-term and large-scale coral-macroalgae regime shifts
[@arif2022; @wild2011]. Climate impacts on coral reefs are decreasing the ability of coral
reefs to provide ecological functions such as benthic habitat and nutrient cycling
[@forster2021; @williams2019] as well as ecosystem services such as food, protection,
tourism and Indigenous cultural value [@eddy2021; @stoeckl2021].

Understanding the characteristics of reefs that will persist under projected climate
conditions requires examining reef resilience across multiple scales and processes. Coral
reef ecosystem resilience - the capacity to maintain coral populations and avoid ecosystem
regime shifts - depends on three interconnected mechanisms. First, the ability of individual
corals to survive thermal stress. Second, the connectivity networks that enable
recolonisation after disturbances [@hock2017; @mumby2008; @torres2024]. Third, the
maintenance, and accretion, of the carbonate framework that provides reef structure
[@perry2012]. We aim to assess how these mechanisms interact under projected climate
conditions to identify the characteristics of resilient reefs through scenario-based coral
population modelling for the Great Barrier Reef ecosystem.

The Great Barrier Reef (GBR) is the world's largest coral reef ecosystem consisting of
thousands of individual reefs. They are interconnected through coral and fish larval
dispersal patterns mediated by hydrodynamic, spatial, and ecological processes [@cappo2000].
The GBR faces severe projected coral cover declines under climate change [@bozec2025;
@bozec2022; @condie2021]. These coral cover declines will be compounded by ocean
acidification, tropical cyclones, and crown-of-thorns starfish outbreaks [@death2012;
@fox-kemper2021; @seneviratne2021; @sweatman2011; @wolff2018]. The 2023-24 Summer mass
bleaching event may be signalling a shift towards this reef future, with some level of
bleaching recorded on 79% of surveyed reefs (total 1001 surveyed reefs)
[@australianinstituteofmarinescience2024; @cantin2024].

At the individual coral scale, survival during thermal stress events depends on both
environmental and biological factors. Environmental factors can either amplify or reduce
thermal stress exposure. Depth is a crucial factor that can provide a refuge from thermal
stress [@álvarez-noriega2018; @baird2018; @smith2014]. Biological factors including coral
size [@brandt2009], taxa [@álvarez-noriega2018; @mcclanahan2005], and thermal tolerance
through acclimation and inheritance [@humanes2022; @lachs2024] all influence survival
outcomes. Importantly, corals demonstrate adaptive capacity to develop thermal tolerance
that could delay widespread bleaching under moderate emission scenarios [@bairos-novak2021;
@humanes2022], representing a crucial mechanism for reef persistence.

Beyond individual survival, reef-scale resilience depends on connectivity networks that
enable recolonization after disturbances [@angeler2023; @elmhirst2009; @hock2017]. However,
connectivity creates both strengths and vulnerabilities. Larval source populations that
export more larvae than they import are essential for system-wide recovery, providing larvae
that enable recolonization at distant locations [@hock2017; @thomas2015]. Yet these source
reefs are paradoxically more vulnerable individually, relying heavily on local recruitment
for their own recovery. When disturbances impact major source populations, cascading effects
reduce larval export and slow recovery throughout the connected network [@figueiredo2022;
@hock2017; @munday2009].

Long-term reef viability requires not only coral survival and recolonization, but also
maintenance of the physical reef structure through positive carbonate budgets. Scleractinian
corals build the primary habitat framework, but reef persistence depends on the balance
between carbonate accumulation [@perry2015] and loss through bioerosion, physical damage
[@kenyon2023], and dissolution [@eyre2014; @eyre2018]. This balance is influenced by coral
cover and community complexity [@perry2012; @desbiens2021], environmental processes, and
ocean chemistry [@perry2015; @eyre2018]. For GBR mid-shelf and offshore reefs, maintaining
positive carbonate budgets requires approximately 17% live coral cover [@desbiens2021].

Recent advances in whole-ecosystem larval connectivity modelling **\[Insert GBR1
connectivity citation.\]** and coral thermal response research [@bairos-novak2021] enable
integrated investigation of how multiple factors interact to determine climate resilience.
By incorporating the latest coral population models [@ribeirodealmeida2024], ecosystem-wide
connectivity data, and updated climate projections [@dixon2022], we can examine which
factors most strongly influence reef persistence under different scenarios. While
connectivity has been emphasized as critical for GBR resilience [@hock2017], and remains an
important consideration, our integrated modelling reveals depth may play a fundamental role
in determining maintenance of reef coral populations and positive carbonate budgets under
current projections of climate change.

# Methods

## Reef definitions

The spatial extent of the GBR reefs represented in this investigation follows the reef
definitions available via the Great Barrier Reef Marine Park Authority (GBRMPA)
[@GBRMPA2022]. The reef definitions define 3806 reefs across the GBR. An indicative map of
the GBR marine park including the Marine Park Management areas [@GBRMPA2007] and the
corresponding bioregion for each reef is shown in (@fig-region-map; [@GBRMPA2024]).
Estimates of coral carrying capacity for each reef were taken from prior analyses undertaken
in [@bozec2022].

Depth characteristics for each reef were extracted using 10m resolution bathymetry data made
available in GBRMPA [-@GBRMPA2020]. While the bathymetry data adopted for the given
assessment are said to be accurate up to 15m from Mean Sea Level [@GBRMPA2020], there are
many sources of uncertainty in the depth estimates used for this study. Notably, the adopted
depth estimates were derived from bathymetries based on satellite imagery. The bathymetry
data can be limited in their alignment with the established GBRMPA reef outlines used to
characterise reef locations and their areal extent. For each reef, bathymetry data were
spatially matched with the 3806 reef outlines through geometric intersections.

A Quality Assessment and Control process identified reefs with potential data limitations:
reefs with no overlapping bathymetry data, reefs with minimum depth values above sea level
(indicating potential data artifacts), and reefs where bathymetry pixels covered less than
5% of the reef polygon area. This process removed 842 reefs, or 22%, from the initial set of
3806 reefs. Furthermore, reefs that are not associated with GBRMPA Management areas
[@GBRMPA2007] or reefal bioregions [@GBRMPA2024] were also removed from the investigation.
Bioregions which contained fewer than three reefs after these processes were also excluded.
As a result, 1377 reefs were excluded (36%), leaving a total of 2429 remaining. All depth
values were converted to positive values (depths below sea level). The median depth was used
to provide an indicative depth value for each reef. Further limitations are outlined in the
Discussion. Code to extract depth estimates, and the resulting depth estimates themselves,
are available at [@iwanaga2025]. To ensure effective data analysis, reefs with incoming or
outgoing connectivity values below the 5th percentile were removed from post-hoc analyses,
leaving 2197 reefs remaining for display in results.

![Map of the Great Barrier Reef region, including GBRMPA Management areas (polygons) and all
considered reefs (points). Colour-coded polygons outline GBR marine park management regions,
where orange represents the Far Northern management region, green represents
Cairns/Cooktown, red covers Townsville/Whitsunday and blue covers Mackay/Capricorn
[@GBRMPA2007]. Reef centroids are marked individually for visual clarity and are coloured
according to its bioregion [@GBRMPA2024]. All 2197 reefs included in this investigation are
represented in this figure. Annotations represent major coastal settlements along the
GBR.](../figs/region_map.png){#fig-region-map fig-align="center" width="14.8cm"}

## Degree Heating Week trajectories

Future projections of maximum yearly Degree Heating Weeks (DHWs) for each reef were
estimated from Sea Surface Temperature (SST) data for the GBR. Statistically downscaled
1km^2^ resolution SST data for five Global Climate Models (GCMs) were obtained from the
HighResCoralStress database [@dixon2022; @dixon2023]. These five GCMs were EC-Earth3-Veg,
ACCESS-CM2, ACCESS-ESM1-5, NorESM2-MM and GFDL-CM4. The selected GCMs are a part of the
Sixth Phase of the Coupled Model Intercomparison Project (CMIP6), spanning four 'model
development families' as defined by Brunner et al., [-@brunner2020].

These GCMs were selected based on evaluation conducted by Grose et al., [-@grose2023],
indicating their ability to recreate historical atmospheric temperature and precipitation
conditions in the Australasian region. They span a range of Equilibrium Climate Sensitivity
(ECS) values within the Intergovernmental Panel on Climate Change (IPCC) assessed likely
range. The IPCC defines ECS as the long-term warming caused by a doubling of carbon dioxide
above its pre-industrial concentration [@forster2021]. While this assessment does not
consider oceanic data, Grose et al. [-@grose2023] provide a region-specific selection
mechanism which we use as an indicator of general performance of GCMs for the region. An
overview of the selected GCMs is provided in @tbl-gcms. Note here that Grose et al.,
[-@grose2023] assessed GFDL-ESM4, whereas Dixon et al., [-@dixon2022; -@dixon2023] provides
data based on GFDL-CM4. This substitution is assumed appropriate given they share the same
fundamental representation of the physical climate system.

Shared Socio-economic Pathway (SSP) 2-4.5 was used to represent the plausible emission
pathway. It represents a moderate emissions scenario with continued CO~2~ emissions
comparable to present day until 2050, after which emissions gradually decline
[@hausfather2020; IPCC -@climate2023]. The selected GCMs represent 2.5°C to 4.66°C of global
atmospheric warming in the 2081-2100 period relative to 1850-1900. The use of these GCMs
therefore allow for a range of possible climate futures to be explored, capturing
uncertainty in future responses to CO~2~ increases (@sfig-ecs; @fig-methods-dhws;
[@grose2023; @shepherd2018]) and enables the assessment of modelled coral reef behaviour
over a broad range of independently derived climate futures [@abramowitz2019].

Monthly mean SST climatological baselines, for each 1km^2^ pixel, were calculated from the
daily SST data following NOAA methodology [@liu2017; @skirving2020]. For each pixel of SST
data, the Monthly Mean SST climatology was calculated by computing the mean SST value for
each calendar month (January through December) across the 28-year baseline period from 1985
to 2012. For each month separately, this climatological baseline was adjusted to centre at
year 1988.2857, which attempts to account for any long-term trends within the baseline years
themselves (see Skirving et al. [@skirving2020]). The adjustment ensures that the
climatology represents conditions at a specific temporal reference point rather than being
influenced by gradual warming or cooling trends that may have occurred during the 1985-2012
period. The resulting monthly climatologies serve as a standardized reference against which
current SST conditions can be compared to identify thermal anomalies and assess coral
bleaching risk, following the methodology established by Skirving et al. [-@skirving2020].

Degree Heating Weeks (DHWs) were derived using the "HotSpot SST Anomaly" approach
[@skirving2020]. The "HotSpot SST Anomaly" represents thermal stress experienced by coral
reefs and is calculated as the difference between daily SST values and the maximum of a
pixel's monthly mean climatology across all twelve months (i.e., the maximum monthly mean
SST). This approach uses the annual maximum temperature as the baseline because corals are
adapted to withstand their local warmest conditions under normal seasonal cycles. When
temperatures exceed this seasonal maximum, corals experience thermal stress that can lead to
bleaching. Only positive anomalies, greater than 1°C, are considered relevant for coral
stress, as temperatures below the seasonal maximum do not contribute to bleaching risk.

DHW values on a per-pixel basis at 1km^2^ resolution were then calculated as the 12-week
rolling sum of HotSpot SST Anomaly values that are greater than 1°C, expressed in units of
°C-weeks. The 12-week accumulation period reflects the cumulative nature of coral thermal
stress. Brief temperature spikes may be tolerable, but sustained or repeated exposure to
elevated temperatures causes progressively greater physiological stress and bleaching
severity. DHW values integrate both the magnitude and duration of thermal stress: a moderate
anomaly (e.g., 1°C) sustained for 12 weeks produces the same DHW value as a severe anomaly
(e.g., 3°C) lasting 4 weeks, though the biological impacts may differ. DHW estimates for
each indicative reef (see @GBRMPA2022) were then determined by taking the mean DHW value for
all pixels touching each reef outline. The process described here was conducted for all 2197
reefs to create DHW projections spanning from 2025 to 2075. Code used to calculate DHW data
and perform analyses involved in this investigation can be found at Grier and Iwanaga
[-@grier2025].

::: {#tbl-gcms}
| Global Climate Model source | ECS value (°C)     | Interpreted relative severity | Results found in |
|-----------------------|-----------------------|------------------------|-----------------------|
| ACCESS-CM2                  | 4.66 [@grose2023]  | High                          | [@grier2025]     |
| EC-Earth3-Veg               | 4.30 [@wyser2020]  | High                          | [@grier2025]     |
| GFDL-CM4                    | 4.1 [@sentman2025] | High                          | [@grier2025]     |
| ACCESS-ESM1-5               | 3.87 [@ziehn2020]  | Intermediate                  | Main text        |
| NorESM2-MM                  | 2.5 [@seland2020]  | Low                           | [@grier2025]     |

Information for each GCM used in this investigation including Equilibrium Climate
Sensitivity (ECS) value, interpreted relative severity and where results for an individual
GCM results are found. GCMs in context with assessed plausible ranges of ECS values are
displayed in [@sfig-ecs]. ACCESS-ESM1-5 has an ECS closest to the midpoint of the
very-likely ECS range and was taken as an intermediate model with results displayed in the
main text.
:::

![Degree Heating Week time series (in °C - Weeks) for the GCMs selected. Colours represent
the different Management areas across the Great Barrier Reef. Green represents
Cairns-Cooktown, orange represents Far-Northern, blue represents Mackay-Capricorn and red
represents the Townsville-Whitsunday Management area. Solid lines represent the
management-area median time series and bands represent the 95th percentile confidence
intervals.](../figs/methods_dhw_timeseries.png){#fig-methods-dhws fig-align="center"
width="14.82cm"}

## Assessment framework

We adapted an exploratory scenario modelling approach for this investigation. Specifically,
we use combinations of model factors to investigate a wide range of potential futures under
different assumptions of environmental (climatic) and ecological conditions. Exploratory
modelling approaches have gained favour in recent years to explore the implications of deep
uncertainty; where there is little or no agreement on the key drivers of a system, the
extent of the relationship, how best to represent them or their probability distributions
[@conradt2024; @lempert2013]. Exploratory modelling embraces uncertainty as an inherent
characteristic of complex environmental systems [@lempert2010; @lempert2013] and is a core
methodology underpinning the development of robust and adaptive management strategies
[@haasnoot2013].

Rather than attempting to forecast precise outcomes, the exploratory approach aims to
systematically search many possible futures to identify patterns that lead to outcomes of
interest. The approach allows for the identification of robust management strategies that
perform well across a diversity of future conditions rather than attempting to identify a
single most optimal strategy. Concurrently, assessing the wide swathe of scenarios allows
contextual understanding of the conditions in which strategies may fail. Multiple
interacting and compounding sources of uncertainty, and a lack of broad agreement as to
their likelihood, create inherently uncertain outcomes. Here, exploratory modelling was
applied to identify the environmental and reef characteristics, such as thermal stress,
depth, and connectivity, that support favourable outcomes in reef recovery and resilience.
Metrics used in analyses can be found in @stbl-metrics.

All data preparation, modelling and analyses were performed using the Julia programming
language (The Julia Project; https://julialang.org/; [@bezanson2017]), with visualisations
created using the Makie ecosystem [@danisch2021].

### Scenario modelling

Two separate systems were employed to explore scenarios: ADRIA (the Adaptive Dynamic Reef
Intervention Algorithms platform) and CoralBlox. ADRIA is a multi-criteria decision support
platform designed for the assessment and analysis of coral reef intervention strategies
under uncertainty [@iwanaga2023]. The ADRIA platform provides functionality for scenario
generation and analysis as well as providing the technical infrastructure to couple
environmental and ecosystem representations. CoralBlox is a computationally efficient coral
population model. It simulates functional/morphological groups of corals over their life
cycles, competing for available space over time [@ribeirodealmeida2024]. The coupled
framework is referred to as ADRIA-CoralBlox.

The coral demographic representation and associated model implementation are further
detailed in the supplementary material. Of important note is that the representation of
bleaching mortalities followed Bozec et al., [-@bozec2022] albeit with a scaling coefficient
to incorporate the amelioration effect that depth may have, following Baird et al.,
[@baird2018]. This depth-scaling reduces the proportion of coral colonies impacted by high
heat stress as depth increases. The impacts of tropical cyclones, wave activity, Crown of
Thorns Starfish or coral disease were not included in this investigation, in order to focus
on assessing reef resilience under possible future warming scenarios.

ADRIA-CoralBlox was run for 4096 scenarios for each GCM (20,480 total model runs). Here, a
scenario represents a unique combination of assumed environmental (namely one of the five
DHW trajectories) and ecological factors including coral growth and mortality rates, and
coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol' sampling
method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a randomised Quasi Monte
Carlo approach which balances scenario exploration with improved convergence properties
compared to pseudo-random sampling. A detailed comparative analysis of sampling methods can
be found in [@kucherenko2015]. For the purpose of scenario assessment, the resulting coral
cover time series for each model run were scaled by the total area of each reef to represent
cover as a percentage of the total estimated reef area. The scenario-median time series for
each reef was used for further analyses. These time series of coral cover were then
clustered based on their temporal behaviour.

### Connectivity modelling

Larval transition probability matrices were generated using the EReefs Great Barrier Reef
1km resolution hydrodynamic model and particle tracking simulations. Connectivity data from
24 days in total were used in this investigation to estimate coral larval flows. These data
span six years from 2015 to 2020. The years 2015, 2016, 2018 and 2019 are each represented
by connectivity matrices covering spawning periods over three days in each year (12 days in
total), while 2017 and 2020 are each represented by six spawning days each, coming to a
total of 24 days of connectivity matrices. The values in these matrices indicate the
probability of larval transfer between each pair of reefs in the GBR, herein referred to as
connectivity strengths. Indicative connectivity strengths are implemented in ADRIA-CoralBlox
as the mean of all available data for each location (i.e., mean of all 24 spawning days).
While the connectivity strengths do not vary from year to year within each ADRIA-CoralBlox
simulation, they do vary between reefs which are of varying sizes, providing sufficient
heterogeneity to capture spatial patterns in larval dispersal and settlement dynamics. This
spatial variability also enables investigations into the influence of connectivity strength
and its role in driving recovery processes.

To assess the impact of reef connectivity on modelled outcomes we summarise incoming
connectivity (weighted incoming connectivity), outgoing connectivity (outgoing connection
strength), the ratio of source-to-sink connectivity, and larval retention probability for
each reef. Weighted incoming connectivity is calculated by summing the product of each
connection strength incoming to a target reef (excluding retained larvae from the target
reef) and the mean timeseries absolute coral cover (km^2^) of each matching source reef.
This metric provides an approximation of overall larval flow into a reef, with high values
indicating the target reef is supplied high volumes of larvae via strong connections and low
values indicating the target reef has weak (or no) incoming connections or is supplied by
small source populations. A limitation of weighted incoming connectivity is it does not
incorporate the proportion of the source reef's population that is reproducing and producing
larvae, instead taking the total coral cover area as an indicator of larval productivity.
Outgoing connection strength is calculated as the sum of all outgoing transition
probabilities from a target reef. Outgoing connection strength has a theoretical maximum of
one (all larvae reach settlement on sink reefs), however the large size of the GBR combined
with larval mortality considerations result in outgoing connection strength values much
lower than one. The ratio of source-to-sink connectivity ($O:I$) for a target reef
represents the strength of the reef's outgoing larval flow relative to it's incoming larval
flow. This is calculated as: $O:I = \frac{S_O/N_O}{S_I/N_I}$. Where S represents the sum of
connectivity pobabilities, N represents the number of connections, O denotes outgoing
connectivity from the target reef and I denotes incoming connectivity to the target reef. If
a reef has no incoming connections, the $O:I$ equals the $S_O/N_O$ value. Incoming
connections in this calculation, do not directly represent the larval supply gained as the
breeding populations from which connections are coming from are not represented. A
source-to-sink ratio greater than 1 (a value greater than 0 on a log scale) indicates a reef
has greater outgoing than incoming larval flow. Reef larval retention probability indicates
the proportion of spawned larvae from a target reef that remain to settlement.

### Time series clustering

Reefs were clustered based on the temporal behaviour of their median coral cover time
series. Put simply, reefs that have similar trends and trajectories are more likely to be
clustered together. Reefs are clustered into Low, Medium, or High, indicating their relative
ability to maintain coral cover over time. Reefs in each time series cluster were compared
to identify characteristics that differ between Low, Medium and High clusters. These
characteristics include reef median depth, mean time series DHW, weighted incoming
connectivity, outgoing connection strength, ratio of source-to-sink larval connectivity,
larval retention probability and reef carrying capacity (coral habitable area).

The Complexity-Invariant Distance (CID) between each time series was calculated to identify
similarities in temporal behaviour [@batista2014]. The CID applies a correction factor to
the Euclidean distance between pairs of time series to obtain a measure of their similarity.
The adopted approach ensures reefs are clustered based on their temporal dynamics as opposed
to, for example, naively clustering by their mean coral cover. Time series clustering is
achieved through the use of k-medoids clustering [cf. @steinmann2020]. Further details on
the time series clustering approach are found in the supplementary materials.

The time series clustering was applied over the time frame between 2025 and 2075. The
clustering was conducted for three spatial scales: (1) the whole of GBR; (2) each GBR
management area; and (3) for each reefal bioregion (defined in GBRMPA [-@GBRMPA2024]). The
analyses was conducted at the different spatial scales as, conceptually, the temporal
behaviour constituting "low", "medium" or "high" coral cover may differ at the different
spatial scales. As it was, analyses at the whole-of-GBR scale were found to lead to
identical conclusions regarding reef characteristics which influence coral cover and so for
brevity are not shown here.

To compare time series clustering results across GCMs, management area scale clustering
results are used to preserve coarse scale latitudinal differences between regions, while
maintaining ease of comparison and visualisation. These comparisons include quantification
of differences between low and high cover clusters across management regions under each GCM
using the CID between cluster median time series. To assess how different reef
characteristics influence time series clustering at a finer spatial scale, bioregion results
are used to capture patterns occurring within individual groups of reefs. Results of all
analyses at all spatial scales can be found at Grier and Iwanaga [-@grier2025] for those
interested.

#### Partial Dependence Analysis

Partial Dependence Analysis was conducted to understand the relationship between features
and bioregion-scale time series cluster predictions using a trained Random Forest (RF)
classifier [@friedman2001]. The analysis utilised the reef characteristics of interest as
the predictors, including the median reef depth, mean DHW for the given GCM, log10-weighted
incoming connectivity, log10-weighted outgoing connectivity, and log10-outgoing-to-incoming
connectivity ratio. The full dataset consisted of the time series cluster assignment of each
of the 2197 reefs under the five GCMs explored, for a total of 10,985 samples. The dataset
\*\*\[\] was randomly split into training (60%) and testing (40%) subsets to ensure model
generalisability. The RF classifier achieved 80% accuracy on the training set and 59% on the
test set.

A 3-fold cross-validation was also conducted, arriving at a similar 59% accuracy rate.
Furthermore, F1 performance scores for Low, Medium, and High clusters were determined to be
0.70, 0.41 and 0.63 respectively, with precision of 0.77, 0.47 and 0.64 when evaluated with
the test set (consisting of 4394 samples). Lower performance with the Medium cluster
reflects the inherent difficulty of classifying transitional states in natural systems. The
consistent level of accuracy across test and cross-validation demonstrates the generalised
performance of the RF classifier, while not accurate, was deemed sufficiently informative
for the purpose of assessing the influence of reef characteristics.

Partial dependence analysis was performed by applying the trained model to the test dataset.
Values for individual predictors were systematically (one-at-a-time) varied across a grid of
50 equally spaced points spanning the predictor's range, while all other features were left
at their modelled values. The procedure creates (1) a modified test dataset, where the
target predictor was set to the grid value for all samples, (2) probabilistic predictions
for all three cluster classes using the trained RF classifier, and (3) the mean predicted
probability for each class across all test observations. Use of the median predicted
probability was also assessed but was found not to substantially influence the
interpretation of results (and so, are not shown here). The process isolates the marginal
effect of each predictor by averaging over the natural distribution of all other
representative reef characteristics in the test dataset. Additionally, to assess possible
interaction between reef depth and incoming connectivity, two-way partial dependence
analysis was performed on grids along the range of these two features.

### Carbonate budget analysis

Using coral cover time series for each reef we calculated the number of years each reef
maintains coral cover above a given percent threshold. The threshold represents the level of
live coral cover necessary to maintain a positive carbonate budget (denoted here as
$\theta_{0}$ to indicate the budget equilibrium). Greater number of years above $\theta_{0}$
suggests greater resilience and opportunity for recovery.

Recent analyses suggest a general threshold of \~17% live coral cover in the case of the
mid-shelf and offshore GBR reefs [@desbiens2021]. Given in situ carbonate budgets are
influenced by (1) a range of processes; (2) the variety of approaches that could be used to
measure those processes; and (3) the uncertainty surrounding these, it would be unreasonable
to assume all reefs adhere to a single precise $\theta_{0}$ value. Concurrently, it is
difficult to ascertain a suitable $\theta_{0}$ value for each individual reef.

Given the uncertainties present, the proportion of modelled reefs that display capacity to
maintain coral cover above a given $\theta_{0}$ threshold is examined. An indicative range
of 10 to 20% is used (inclusive, with a step size of 1%). The range was informed by Perry et
al., [-@perry2015] and Desbians et al., [-@desbiens2021] providing plausible indications of
the lower and upper bounds respectively. Robustness of the conclusions drawn to the
uncertainty in the precise value of $\theta_{0}$ for all reefs can then be assessed by using
the indicative range.

The relationships between the number of years above $\theta_{0}$ and characteristics such as
median reef depth and Log-10 total connectivity strength were then assessed with Spearman's
rank correlation. Spearman's correlation coefficient was adopted as reef depth and
connectivity strength distributions are expected to be non-normal and highly skewed.
Furthermore, reefs are naturally of different types and experience different conditions, and
so the monotonic trends (whether the relationship consistently increase or decrease) rather
than the exact quantitative relationship was of interest for this analysis.

# Results

## Reef trajectories across futures

There were substantial differences in the maintenance of coral cover post 2030 between the
GCMs. Figure [-@fig-gcm-cover-trajectories] illustrates the temporal behaviour of reefs
across clusters (Low, Medium and High).

Of the GCMs explored, NorESM2-MM had the lowest Equilibrium Climate Sensitivity (2.5°C) and
reefs exhibit continuing recovery and decline dynamics up to 2075 under this hypothetical
climate future. Unsurprisingly, recovery phases were predominantly seen in the reefs in the
High cluster, with Medium and Low exhibiting lesser probability of such periods of recovery.
Generally, recovery and increases in coral cover were visible in simulated results up to
2045. Under NorESM2-MM, the Far Northern Management area were found to have the most similar
temporal behaviour between Low and High clusters compared to other Management areas
(@stbl-cluster-cid) due to the low degree of coral cover overall
(@fig-gcm-cover-trajectories).

Intermediate levels of coral cover maintenance can be seen in ACCESS-ESM1-5 (ECS value of
3.87°C) when compared to the other GCMs (of higher and lower ECS values). Under these
intermediate conditions (in relative terms), there is potential for reefs in the High
cluster to maintain coral cover levels around 20% to 2050. The highest peak DHW levels occur
under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs time series
occur under ACCESS-CM2 (@fig-methods-dhws). Under conditions represented by high ECS value
GCMs - specifically, ACCESS-CM2 (4.66°C), EC-Earth3-Veg (4.30°C) and GFDL-CM4 (4.1°C) - all
reefs are largely in decline across all management areas from 2055 (@stbl-cluster-cid).

![Time series of coral cover (relative to reef area) categorised within each Management area
(rows), for each GCM analysed (columns) based on temporal behaviour over 2030 and 2060. The
time series are categorised as "Low" (green), "Medium" (in orange), and "High" median cover
(in blue). Solid lines represent the cluster median time series and bands represent the 95th
percentile confidence
intervals.](../figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
fig-align="center" width="14.82cm"}

Comparing across GCMs, it is common for reefs to switch between clusters, including from Low
to High and vice versa (@sfig-gcm-bioregion-cluster-assignment). Of 2197 reefs, 21% (462) of
reefs have consistent cluster assignment across all GCMs. Of the 462 consistently assigned
reefs, 23% are within the Far Northern Management area, with 1%, 55% and 21% occurring in
the Cairns-Cooktown, Townsville-Whitsunday and Mackay-Capricorn Management areas
respectively (@sfig-gcm-cluster-map). Of the 462 consistently assigned reefs, 34%, 18% and
47% are clustered into Low, Medium and High clusters respectively. High levels of assignment
variability across GCMs are expected as due to spatial differences in each given GCM’s DHW
trajectories.

### Bioregion scale clustering

In contrast to the broad temporal behaviour at the Management area scale, the relationship
between reef characteristics and assigned clusters were found to be similar at the bioregion
scale across GCMs. Results are only shown and discussed for ACCESS-ESM1-5, an intermediate
ECS scenario (@sfig-ecs). Those interested in results for the other assessed GCMs are
referred to Grier and Iwanaga [-@grier2025]. Across bioregions, reefs in the High cluster
not only had greater levels of coral recovery but displayed persistence to 2050
(@sfig-access-esm1-5-bioregion-timeseries). Reefs in the High cluster were consistently
found to have the greatest median reef depths across bioregions
(@fig-access-esm1-5-bioregion-depth).

![Median depth for reefs within each bioregion, separated into clusters assigned by time
series clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative
cover between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters
with intermediate median area-relative cover between 2030 and 2060 within each bioregion.
‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030
and 2060 within each bioregion. Solid lines represent the median depth for each cluster
across reefs and bands represent the distribution of depth values within each cluster.
Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/depth_bioregion_violin.png){#fig-access-esm1-5-bioregion-depth
fig-align="center" width="14.82cm"}

Across bioregions there were no consistent patterns between reef clusters in mean DHW values
(@sfig-access-esm1-5-bioregion-dhwmean), Log-10 weighted incoming connectivity
(@sfig-access-esm1-5-bioregion-weightedincoming), Log-10 outgoing connection strength
(@sfig-access-esm1-5-bioregion-outgoingconn), Log-10 source-to-sink connectivity ratio
(@sfig-access-esm1-5-bioregion-logsosi) and Log-10 larval retention probability
(@sfig-access-esm1-5-bioregion-selfconn). Across six of the 29 bioregions reefs in the high
cover cluster have lower Log-10 carrying capacity values than reefs in other clusters
(@sfig-access-esm1-5-bioregion-carrying-capacity).

#### Partial dependence

A Partial Dependence Analysis was conducted using a Random Forest classifier to identify
reef characteristics that influence its temporal behaviour, and thus its cluster assignment
into Low, Medium, or High. Reef median depth, carrying capacity and mean time series DHW
level have the highest feature importance scores, accounting for 29%, 15% and 15% predictive
importance respectively (@fig-ts-cluster-pdp; Panel A). Reef weighted incoming connectivity,
outgoing connectivity strength, source-to-sink connectivity ratio and larval retention
probability account for \~12%, 11%, 10% and 8% respectively.

Reefs have the greatest probability of being clustered into Low if shallow, and High at
greater depths, with medium cluster reefs exhibiting a lower, and more stable, assignment
probability across depth ranges (@fig-ts-cluster-pdp; Panel B). A large increase in the
probability of a reef being categorised as High can be seen between 10 and 15 meters depth.
This is at the edge of the depth range used in current deployment efforts [@howlett2021;
@randall2021; @whitman2024] but indicates a depth range that could be adopted to maximise
longevity and benefits of, for example, coral seeding activities. Across mean DHW values,
assignment probability is roughly equal across the three clusters up to 5 DHW. A sharp
delineation can be seen \> 5 DHW when comparing High with other clusters, with lower
assignment probability (\~25%) compared to Medium and Low (\~40%) (@fig-ts-cluster-pdp;
Panel B).

![Feature importance scores (Panel A) of the Random Forest classifier, and Partial
Dependence Plots (PDPs) for reef characteristics of interest indicating the probability of a
reef being classified to one of the three cover clusters (Panel B). Partial Dependence Plots
for bioregions were excluded as they were found to be uninformative. PDPs indicate the
probability of a reef from the test dataset being clustered into each cover cluster
depending on the predictor value on the x axis, when adjusted for values of other
predictors. Green, orange and blue lines indicate cluster assignment probability for Low,
Medium and High clusters respectively. Transparent grey points above probability lines
represent the predictor values present in the test dataset, providing a rough indication of
predictor value distribution across
reefs.](../figs/ts_cluster_rf_feature_analysis.png){#fig-ts-cluster-pdp fig-align="center"
width="14.82cm"}

Analyses for source-to-sink ratio and incoming and outgoing connectivity metrics suggest
that reefs with greater outgoing connectivity and lower incoming connectivity levels have
higher probability of being clustered into the Low cover cluster compared to other reefs.
Additionally, two-way analyses revealed little interaction between depth and incoming and
outgoing connectivity metrics for reef clustering probabilities
(@fig-ts-cluster-pdp-twoway-incoming; @sfig-ts-cluster-pdp-twoway-outgoing). Reefs with a
median depth around 15m and high incoming connectivity may benefit slightly, with higher
probability of High cover cluster assignment compared to 15m - depth reefs with lower
connectivity.

![Partial dependence plot indicating the probability of cluster assignment along interacting
values of median reef depth (metres) and log10-weighted incoming connectivity. Colour
indicates the probability of assignment to each cluster, with lighter colours indicating
higher probability of assignment. Panel A shows the probability of assignment for the Low
cover cluster, Panel B shows Medium cover cluster and Panel C shows assignment probability
for the High cover cluster. Probability values for each cluster are generated by the random
forest classifier using each combination of depth and incoming connectivity values along the
axes at 50 grid points. The actual distributions of depth and incoming connectivity values
for reefs are displayed in panels D and E
respectively.](../figs/ts_cluster_rf_twoway_depth_incomingconn_pdp.png){#fig-ts-cluster-pdp-twoway-incoming
fig-align="center" width="14.82cm"}

Individual feature analysis for carrying capacity indicates that its influence on cluster
assignment is concentrated at the low range of predictor values (@fig-ts-cluster-pdp). This
aligns with the right-skewed distribution of carrying capacity across reefs for these three
predictor variables. Beyond this low-end range, the partial dependence becomes near
constant, indicating that outlier values for these factors are not influential.

Assessing random forest prediction accuracy across bioregions, reveals increasing
performance as bioregion mean reef depth increases (@sfig-rf-predictor-perf). This
highlights that unexplained variation in the test dataset (33% of variation) may be
concentrated in shallow bioregions where the available model predictors do not capture
differences between clustered reefs.

## Reef carbonate budget

Reef median depth had a moderate positive correlation with the number of years a reef
maintains coral cover above the assumed carbonate budget threshold required to sustain live
coral cover ($\theta_{0}$). Greater median depth was a key characteristic defining reefs
that sustain coral cover above $\theta_{0}$ (@fig-access-esm1-5-depth-conn-carbonate;
@sfig-access-esm1-5-mean-carb-budget; @sfig-access-esm1-5-carb-budget-corr). Unsurprisingly,
the number of years above $\theta_0$ substantially decreased as $\theta_0$ is increased, as
does the Spearman's correlation (@sfig-access-esm1-5-carb-budget-corr). Decreasing
correlation levels are likely due to an increasing number of reefs being unable to maintain
coral cover above $\theta_0$ as it is increased, with the distribution of years above
threshold becoming increasingly right-skewed (@fig-access-esm1-5-depth-conn-carbonate). In
contrast, weighted incoming connectivity strength (Log-10 scale) was found to have no
correlation with the number of years a reef maintains cover above the required thresholds
(@sfig-access-esm1-5-carb-budget-corr).

![Number of years reef coral cover exceeds the carbonate budget threshold from 2025 to 2075
(y-axis) and reef median depth (left panel) and reef log10 weighted incoming connectivity
values (right panel). Point colour represents the threshold assessment values used to assess
positive carbonate budget status ($\theta_0$). Deeper reefs are better able to maintain
cover above the indicative carbonate budget threshold, while there is no considerable
relationship between weighted incoming connectivity and carbonate budget maintenance.
Displays results from median time series using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/depth_conn_carbonate_comparison.png){#fig-access-esm1-5-depth-conn-carbonate
fig-align="center" width="14.82cm"}

# Discussion

Our results show increasing reef depth has a strong positive impact on modelled reef
resilience in temporal behaviour and maintenance of positive carbonate budgets under
possible future conditions. Reefs with a median depth of 10m or greater, with high incoming
connectivity, are the most likely to be assigned to a high cover time series cluster. As an
individual feature, incoming connectivity alone does not correlate with maintenance of
positive carbonate budgets. These findings highlight the importance of depth and incoming
connectivity as coupled selection features when assessing potential reef resilience under
future thermal stress.

Reefs in different temporal clusters largely do not differ in mean modelled DHW levels
experienced and exhibit overall similar DHW time series. While reefs experience similar
magnitude DHWs based on SST, the proportion of a reefs' coral population impacted by heat
stress events depends on reef depth. Reefs at deeper depths are likely to be clustered into
High cover clusters. Reefs at shallow depths experience no, or few, years with coral cover
exceeding levels likely required to maintain positive carbonate budgets. Movement of shallow
reefs towards negative carbonate budgets in the future may contribute to phase shifts in
reef dominant CaCO~3~ production. Shifting from coral driven to encrusting algae driven
production, or to bioerosion dominated systems with rapid erosion of physical reef
frameworks [@dudgeon2010; @perry2008]. Reefs at increased depths may be able to maintain
coral driven CaCO~3~ production and accumulation into the future for longer. Our results
suggest that within shallow-water reef ecosystems (non-mesophotic, depth \< 30m), depth
contributes greatly to reef resilience and the ability of reefs to maintain positive
carbonate budgets under future conditions. Spatially our results indicate that regions of
positive carbonate budget maintenance, under ACCESS-ESM1-5, and high depth levels were
concentrated in the Far Northern, offshore Townsville/Whitsunday, and Mackay/Capricorn
management areas (@sfig-carb-budget-map; @sfig-depth-map). Resilient reefs in the Far
Northern management area were exposed to lower mean DHW levels than other reefs under
ACCESS-ESM1-5 (@sfig-mean-dhw-map), however similar spatial patterns were observed across
GCMs for reefs clustered into the high cover cluster (@sfig-gcm-cluster-map). We do not
interpret these spatial findings as accurate predictions of future resilience due to high
uncertainty in input data, particularly depth and connectivity. While similar findings have
been demonstrated in the Pacific [@smith2014; @baird2018] and Indo-Pacific [@bridge2014]
regions, depth refugia analyses have often focussed on the potential of mesophotic reefs (\>
30m deep) as thermal refugia [@bongaerts2010; @holstein2015; @semmler2017]. In contrast,
Venegas et al. [-@venegas2019] found no association between depth and the alleviation of
heat stress across a range of shallow reefs in the Pacific.

Our findings support the hypothesis that reefs at greater depths will exhibit
characteristics of refugia and more prolonged resilience under climate change compared to
shallower reefs. Additionally, Liu, Callaghan and Baldock [-@liu2024] suggest that rubble
instability is inversely correlated with reef depth, possibly limiting juvenile coral
recruitment capacity of shallow reefs. A caveat to our depth-resilience findings is the
quality of the bathymetry data used to calculate median reef depth. As detailed previously,
estimates of median reef depth are derived from satellite imagery (at $10m^2$ resolution)
which can be subject to high degrees of uncertainty, both in terms of the estimated depth
for each pixel as well as their alignment to the established GBRMPA reef outlines.
Additionally, we do not incorporate light-based decreases in coral growth rates as depth
increases [@baker1975; @huston1985; @ruiz-diaz2022].

Current reef restoration and intervention efforts on the GBR, such as deployment of coral
seeding, coral fragment devices and Crown of Thorns Starfish control, target reefs with
sites in the 1-10m depth range [@howlett2021; @randall2021; @whitman2024]. While there are
no explicit justifications for this depth range, intervention depth is likely constrained by
snorkelling and diving deployment and monitoring logistical accessibility, visual
identification of sites from the surface, in combination with tourism return value. Our
results suggest that of the assessed reefs occurring within the 1-10m depth range (1551
reefs), 59% (920) of reefs are consistently clustered into the Low and Medium cover clusters
across the GCMs (at the GBRMPA management region scale), with many unable to maintain
positive carbonate budgets under future conditions. Partial dependence analysis revealed a
range of depths from 10-15m where the probability of a reef being clustered into the High
cover cluster increased substantially. Additionally, two-way partial dependence analysis
revealed that high incoming connectivity is important for High cluster assignment for reefs
at depths of \>10m. Considering this 10-15m depth range when assessing possible intervention
sites may reduce the exposure of planted and seeded corals to high thermal stress, while
still maintaining effective logistical access to sites. Managers are also likely to consider
the potential for positive flow on effects due to outgoing connectivity when selecting
intervention sites [@hock2017]. Transient, and difficult to measure, properties such as reef
connectivity are likely to increase uncertainty in intervention planning to a greater degree
than environmental characteristics such as reef depth. Overall, reef intervention planning
is a complex and challenging decision-making process [@anthony2020], and our results suggest
that intervention site depth may be an important factor to additionally consider when
assessing the long-term resilience of sites.

It is a common conclusion in coral reef ecology that a primary facet of reef resilience is
the supply of recruits that a reef receives following disturbance [@hock2017; @mumby2008;
@torres2024]. Our results suggest that incoming connectivity contributes positively to high
cover cluster allocation for reefs around 15m deep. However, our results suggest no direct,
first order, relationship between weighted incoming connectivity and the ability of reefs to
maintain positive carbonate budgets under future conditions. There has been contrasting
evidence [@bode2018; @hock2017] surrounding whether recent generation connectivity modelling
is able to identify reefs that are important for recovery processes, with findings dependent
on the larval connectivity model used. These investigations were performed using prior
generations of larval connectivity models parameterised for 2008-2013 [@hock2017] and
1996-2002 [@bode2018] time periods, without considering the influence of reef depth on
resilience. Using updated 1km^2^ resolution larval connectivity across the whole GBR, with a
limited temporal sample of particle flows, our results suggest that larval connectivity
alone may not be important or reef resilience and recovery, but should be considered in
combination with more influential factors such as depth. Comprehensive comparisons and
validation of multiple connectivity models, as well as use of a model ensemble, would
improve certainty and allow for the targetting of reefs by combining abiotic factors such as
depth with connectivity.

Our findings on the importance of connectivity likely differ from literature due to the
highly uncertain nature of larval connectivity modelling [@choukroun2025;
@vogt-vincent2023], the way larval flows and juvenile recruitment are implemented in ADRIA -
CoralBlox, and the magnitude of early bleaching events decreasing the breeding capacity of
strongly connected reefs. Analysing daily connectivity data Vogt-Vincent et al.
[-@vogt-vincent2023] suggest that time-mean connectivity may not be suitable for capturing
the seasonal - ecological time scale connectivity between locations due to high daily
variability in connectivity data. This is supported by the connectivity data used in our
investigation, which have large relative standard deviations for connectivity between reefs
over the 24 days used (@sfig-connectivity-rsd). Using alternative summarisation methods that
capture ecological-timescale patterns in coral larval dispersal, such as cumulated implicit
connectivity [@vogt-vincent2023], future investigations may find increased importance of
both depth and reef connectivity for resilience, in line with other investigations
[@hock2017; @torres2024]. Choukroun et al [-@choukroun2025] performed a comparison of three
GBR-scale connectivity models, including GBR1, and suggest the use of a model ensemble
approach to estimate connectivity metrics due to the high sensitivity of connectivity to
model specification. Further improvements and research in connectivity modelling are
necessary when incorporating connectivity as a supplementary factor in reef resilience and
intervention-planning analyses.

Under all GCMs many reefs are projected to suffer substantial declines in coral cover
relative to 2025 levels, with an average of 38% of reefs (924 reefs) clustered into the Low
cover cluster at the management area scale across GCMs. NorESM2-MM and ACCESS-ESM1-5 GCMs
convey the most coral cover maintenance into the future. Differences between Low, Medium and
High reef clusters in cover time series are most notable under these two GCMs
(@stbl-cluster-cid; @fig-gcm-cover-trajectories). Maintenance of coral cover into the future
is linked to the Equilibrium Climate Sensitivity of the GCM used as levels of future
recovery and timespan of cover maintenance increase as GCM ECS increases
(@fig-gcm-cover-trajectories). Under high-ECS GCMs (GFDL-CM4, EC-Earth3-Veg and ACCESS-CM2)
large declines and lower rates of recovery in high cover clusters are common compared to
lower-ECS GCMs.

The large differences in cover time series seen across the GCMs suggest that it is important
to consider uncertainty in future conditions when assessing possible biological responses to
change. This is supported by recent literature suggesting an effective approach to capturing
climate change uncertainty is to model responses using GCMs separately [@shepherd2018].
Despite large differences in coral cover across GCMs, the patterns identified between low,
medium and high clusters of reefs suggest that the relative importance of environmental and
biological factors such as depth refugia, connectivity and surface DHW levels is consistent
across scenarios.

## Limitations

There are multiple caveats in how coral ecosystem dynamics are modelled in this
investigation. We model coral colonies as circular diameter values that are tracked over
timesteps through growth. Additionally, we do not model the direct influence of light
intensity on coral growth or the negative influence of depth on coral growth rates due to
reduced available light levels [@baker1975; @huston1985; @ruiz-diaz2022]. If included in
this investigation it is possible that reef recovery ability, via regrowth, would be reduced
at greater depths, reducing the positive effect of depth refugia for temporal resilience.
Additionally, reduced growth rates at increased depths may reduce the carbonate budget
maintenance of reefs due to lower rates of CaCO~3~ accretion. Due to the availability of
quality future projected data we do not consider other important impacts such as waves,
cyclones, currents and Crown of Thorns Starfish (CoTS) on coral colony survival and
bleaching [@massel1993; @deaker2022].

Uncertainty in connectivity modelling outputs, satellite-derived bathymetry data and future
thermal stress projections are likely to influence the results obtained in this
investigation. While the resolution of connectivity data used in this investigation improves
upon data used in comparable studies [@bozec2025], the time-mean implementation used may not
effectively capture long term patterns [@vogt-vincent2023]. Using alternative connectivity
implementation methods that better capture the ecological-timescale trends in reef
connections rather than the high daily variability, would improve confidence in the
connectivity findings of this investigation [@vogt-vincent2023]. Additionally greater year -
sample sizes for baseline connectivity data as future computation improvements are made,
along with the derivation of connectivity from an ensemble of biophysical models, will
improve confidence in connectivity data for ecological modelling and management applications
[@choukroun2025]. The DHW data used in this investigation have not been bias corrected for
the Great Barrier Reef region and are part of a wider Australian dataset [@dixon2022]
developed using statistical downscaling. Additionally, Lachs et al. [@lachs2021] suggest the
utility of DHW levels for prediction of coral bleaching can be improved by reducing the
threshold for HotSpot Anomaly values from MMM +1 to MMM and reducing the accumulation window
from 12 weeks to 8 weeks. We investigate a range of future climate responses, thermal
projections and ecological parameterisations of ADRIA - CoralBlox to account for many
possible futures, and our results suggest the factors that influence reef resilience are
consistent across these scenarios.

Additionally, it is currently not possible to obtain highly accurate and precise estimates
of reef depth at the GBR-scale. Data adopted for the given assessment are said to be
accurate up to 15m from Mean Sea Level [@GBRMPA2020; @li2021], however there are many
sources of uncertainty in the depth estimates used for this study. Satellite-derived
estimates are subject to different bottom types (rock, sand, seagrass, etc) and their effect
on reflectance, the water penetration depth of satellite sensors, influence of local
conditions (e.g., turbidity and wave activity), among many other factors. The median depths
for each indicative reef area were extracted from bathymetry data using zonal statistics and
depth values beyond the stated 15m accuracy threshold [@GBRMPA2020] were included. The
resulting depth estimates are therefore highly uncertain and are only adopted as a useful
indication for the purpose of the presented analysis. They are not intended to be an
accurate and precise report of reef depths across the GBR.

In addition to the usual challenges of deriving data from satellite imagery, it is
acknowledged that there is high potential for misalignment between the bathymetry data and
GBRMPA reef outlines. Hammerton and Lawrey [-@hammerton2025], for example, identified 120
false positives, where features identified as reefs were sand banks or rocky reefs (rather
than coral reefs). Reefs with bathymetry data of less than 5% of its indicative area were
excluded from analysis to ensure analyses were not conducted on reefs with no data coverage
(see @sfig-depth-cutoff). The existing GBRMPA reef outlines were adopted to align with other
ongoing modelling and analyses conducted as part of the Reef Restoration and Adaptation
Program (<https://gbrrestoration.org>; [@thereef]).

The modelling conducted with ADRIA-CoralBlox itself is subject to many uncertainties. For
one, a formal mechanistic and spatial validation of the model represents an area for future
refinement and is the focus of a future publication. Indicative results are included in the
supplementary **\[TI: For Pedro to include! :`D`\].** The consistency of projected patterns
with known coral-climate relationships and reef characteristics provides confidence in the
model's ability to meaningfully capture ecological signals at the GBR-scale. The mechanistic
basis of the model, combined with qualitative assessment of model behavior, supports the
robustness of our findings for understanding broad-scale coral cover responses to climatic
drivers. The aim here, however, was to assess the relative influence of reef characteristics
on reef resilience under future conditions, rather than provide spatially explicit and
precise predictions of reef outcomes. Using this approach, we suggest that reef depth is an
important factor for reef resilience under thermal stress.

Improvements in remote sensing approaches for broad scale bathymetry measurement and
understanding of how depth impacts experienced thermal stress on reefs will further enhance
investigations of depth as a critical factor for reef resilience. Additional analyses could
build on the findings of this study by assessing the impacts of prioritising reef depth in
reef intervention decision making. Future analyses could incorporate other environmental and
ecological stressors such as wave activity, ocean acidification, and Crown-of-Thorn Starfish
[@lin2023; @mcclanahan2005; @grimaldi2023; @grimaldi2022; @liu2024; @eyre2018].

## Conclusion

The results presented suggest that when modelled as part of a large and complex ecosystem,
reef depth is a key factor contributing to ongoing reef resilience. We provide evidence in
support of depth as a form of thermal refugia for shallow-water reefs on the GBR. Targetting
reefs with median depths greater than 10m may aid in identifying the most resilient reefs
under climate change. We suggest further analyses that investigate the benefits of
incorporating depth-based resilience into reef intervention decision making, as well as
further research and development in connectivity modelling methods for whole-ecosystem
scales.

# References

::: {#refs}
:::

# Author Contributions

**Benjamin Grier:** conceptualisation, data curation, formal analysis, investigation,
methodology, visualization, writing – original draft, writing – review and editing. **Takuya
Iwanaga:** conceptualisation, data curation, formal analysis, investigation, methodology,
visualization, writing – original draft, writing – review and editing. **\[BG: To be
completed.\]**

# Acknowledgements

We acknowledge the contributions of Barbara Robson and Jessica Benthuysen to manuscript
preparation. This research was funded by the Reef Restoration and Adaptation Program, and
facilitated by the Australian Institute of Marine Science. We acknowledge thermal projection
data sourced from HighResCoralStress datasets, model parameterisation data sourced from the
Ecological Intelligence for Reef Restoration (EcoRRAP) subprogram, larval transition
probability data derived from the eReefs GBR1 hydrodynamic model, and spatial and depth data
available from the Great Barrier Reef Marine Park Authority.

# Conflict of Interest Statement

The authors declare no conflicts of interest.

# Data Availability Statement

The full reproducible investigation, and links to required data sources, can be found at
Grier and Iwanaga [-@grier2025].

# Supplementary Material

![Equilibrium Climate Sensitivity (ECS) values in °C for the GCMs selected (point values).
The pink - dashed box represents the high-certainty ‘very likely’ range for ECS values and
the solid red box represents low-certainty ‘likely’ range for ECS values (Forster et al.,
2021).](../figs/ecs_plot.png){#sfig-ecs fig-align="center" width="10cm"}

![Distribution of GBRMPA [-@GBRMPA2020] satellite derived depth data coverage levels across
3806 possible reefs. Reefs below the 5% data coverage threshold (represented by the red
vertical line) were removed from the
investigation.](../figs/reef_depth_included_cutoff.png){#sfig-depth-cutoff
fig-align="center" width="14.8cm" height="9.5cm"}

![Reef Degree Heating Week time series clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2060 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2060 within each bioregion. Solid lines represent the cluster median time series and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../figs/GCM_dhw_timeseries_plot.png){#sfig-gcm-dhw-trajectories
fig-align="center" width="14.82cm"}

::: {#stbl-metrics}
| Metric | Used in | Definition | Justification |
|-----------------------|-----------------------|-----------------------|-----------------------|
| Percentage coral cover time series | Time series clustering, Carbonate budget analysis | Total coral cover as a percentage of reef area | To summarise coral population health over time. |
| Median depth | Time series cluster assessment Carbonate budget analysis | Reef median depth as calculated with zonal raster analysis | Depth is a critical form of environmental refugia from high thermal stress |
| Mean time series DHW | Time series cluster assessment | Reef mean DHW value across time series | To quantify the surface thermal stress experienced on each reef over time series. |
| Source - to - sink connectivity ratio | Time series cluster assessment | Ratio of reef outgoing to incoming connectivity | To examine whether a reef may be a 'net contributor' or 'net receiver' of larvae in the connectivity network. |
| Weighted incoming connectivity | Time series cluster assessment, Carbonate budget analysis | Incoming connectivity strength weighted by mean time series coral cover of source reefs | To approximate the larval flow received by a reef over the projected time series. |
| Outgoing connectivity strength | Time series cluster assessment, Carbonate budget analysis | Sum of a reef's outgoing larval transition probability values. | To understand the contribution of a reef to the connectivity network and the possibility for flow-on effects from reef states. |
| Larval retention probability | Time series cluster assessment | Diagonal larval transition matrix values. | To understand the relative proportion of larvae retained on a reef. |
| Carrying capacity | Time series cluster assessment | Reef habitable area \[m^2^\] | Captures the size of the viable reef extent. |

Quantitative metrics used in this investigation.
:::

```{julia}
#| echo: false
#| warning: false
#| message: false
#| output: asis

cap = """
Complexity invariant distance (CID) between low and high cluster median time series in each management area under each GCM. CID values are calculated for percentage cover time series (@fig-gcm-cover-trajectories) and Degree Heating Week (DHW) time series (@sfig-gcm-dhw-trajectories). Complexity invariant distance represents the Euclidean distance between time series, corrected by the complexity of the two time series. CID values are used to quantify overall differences between these two cluster levels at management area and GCM levels. CID values for percentage cover and DHW time series are not directly comparable as the time series use different units.
"""
using Markdown,Tables,MarkdownTables
cd("../src")
include("../src/ADRIA/HighResCoralStress/6_quantifying_cluster_differences.jl"); # Quantify low-high cluster differences using CID metric.
rename!(distances,
    "management_area" => "Management area",
    "CID_cover" => "CID - cover",
    "CID_dhw" => "CID - DHW"
)
print("""
::: {.table #stbl-cluster-cid}
$(markdown_table(distances, String))

$(cap)
:::
""")
```

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by management area scale cluster assignment.
Colour-coded areas outline GBR marine park management regions, where orange represents the
Far Northern management region, green represents Cairns/Cooktown, red covers
Townsville/Whitsunday and blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are
marked individually for visual clarity. Reefs that do not have a consistent cluster
assignment level across all GCMs at mangement area scale are coloured grey. All consistently
clustered reefs are coloured according to their management area scale cluster assignment.
Green represents reefs in the low cover cluster, orange represents reefs in the medium cover
cluster and blue represents reefs in the high cover cluster. All 2197 reefs included in this
investigation are represented in this figure. Annotations represent major coastal
settlements along the GBR.](../figs/gcm_cluster_assignment_map.png){#sfig-gcm-cluster-map
fig-align="center" width="14.82cm"}

![Heatmaps indicating reef cluster assignments within each bioregion across the five GCMs.
Green indicates a low cover cluster assignment, orange represents a medium cover assignment
and blue represents assignment of a reef to a high coral cover cluster. The vertical width
of reef elements varies depending on the number of reefs within the bioregion (displayed on
the y axis). The x axis represents the different GCM levels (EC-Earth3-Veg, ACCESS-ESM1-5,
ACCESS-CM2, NorESM2-MM and
GFDL-CM4).](../figs/GCM_bioregion_cluster_assignments.png){#sfig-gcm-bioregion-cluster-assignment
fig-align="center" width="14.82cm"}

![Partial dependence plot indicating the probability of cluster assignment along interacting
values of median reef depth (meters) and log10 weighted outgoing connectivity. Colour
indicates the probability of assignment to each cluster, with lighter colours indicating
higher probability of assignment. Panel A shows the probability of assignment for the Low
cover cluster, Panel B shows Medium cover cluster and Panel C shows assignment probability
for the High cover cluster. Probability values for each cluster are generated by the random
forest classifier using each combination of depth and weighted outgoing connectivity values
along the axes at 50 grid points. The actual distributions of depth and outgoing
connectivity values for reefs are displayed in panels D and E
respectively.](../figs/ts_cluster_rf_twoway_depth_outgoingconn_pdp.png){#sfig-ts-cluster-pdp-twoway-outgoing
fig-align="center" width="14.82cm"}

![Scatter plots indicating the bioregion test-prediction accuracy (proportion of reefs in
each bioregion assigned correct clusters) from the Random Forest model. Plot panels
represent the bioregion accuracy across various predictor ranges and bioregion attributes,
such as mean bioregion depth, carrying capacity, DHW value, total connectivity strength,
Source-to-sink ratio and the number of reefs in each
bioregion.](../figs/rf_performance_across_predictors.png){#sfig-rf-predictor-perf
fig-align="center" width="14.82cm"}

![Heatmap indicating reef connectivity relative standard deviation (RSD %). Reef
connectivity relative standard deviation represents the reef standard deviation connection
strength as a percentage of reef mean connection strength. Lighter colours indicate higher
RSD % values.](../figs/connectivity_rsd_percent.png){#sfig-connectivity-rsd
fig-align="center" width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by median reef depth. Colour-coded areas outline GBR
marine park management regions, where orange represents the Far Northern management region,
green represents Cairns/Cooktown, red covers Townsville/Whitsunday and blue covers
Mackay/Capricorn [@GBRMPA2007]. Reef centroids are marked individually for visual clarity.
Darker colours represent greater reef depth values than lighter colours. All 2197 reefs
included in this investigation are represented in this figure. Annotations represent major
coastal settlements along the GBR.](../figs/depth_med_input_map.png){#sfig-depth-map
fig-align="center" width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by reef carrying capacity (coral habitable area).
Colour-coded areas outline GBR marine park management regions, where orange represents the
Far Northern management region, green represents Cairns/Cooktown, red covers
Townsville/Whitsunday and blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are
marked individually for visual clarity. Darker colours represent greater carrying capacity
area values than lighter colours. All 2197 reefs included in this investigation are
represented in this figure. Annotations represent major coastal settlements along the
GBR.](../figs/abs_k_area_input_map.png){#sfig-abs-k-area-map fig-align="center"
width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by outgoing connection strength (log10). Colour-coded
areas outline GBR marine park management regions, where orange represents the Far Northern
management region, green represents Cairns/Cooktown, red covers Townsville/Whitsunday and
blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are marked individually for
visual clarity. Darker colours represent higher outgoing connection strength values than
lighter colours. All 2197 reefs included in this investigation are represented in this
figure. Annotations represent major coastal settlements along the
GBR.](../figs/log_out_strength_input_map.png){#sfig-outgoing-conn-map fig-align="center"
width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by source-to-sink connectivity ratio (log10).
Colour-coded areas outline GBR marine park management regions, where orange represents the
Far Northern management region, green represents Cairns/Cooktown, red covers
Townsville/Whitsunday and blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are
marked individually for visual clarity. Darker colours represent higher source-to-sink
values than lighter colours. Source-to-sink ratio values above zero (one on normal scale)
indicate a reef has greater outgoing than incoming connectivity. All 2197 reefs included in
this investigation are represented in this figure. Annotations represent major coastal
settlements along the GBR.](../figs/log_so_to_si_input_map.png){#sfig-sotosi-conn-map
fig-align="center" width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by larval retention probability (log10). Colour-coded
areas outline GBR marine park management regions, where orange represents the Far Northern
management region, green represents Cairns/Cooktown, red covers Townsville/Whitsunday and
blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are marked individually for
visual clarity. Darker colours represent higher probability of retaining spawned larvae than
lighter colours. All 2197 reefs included in this investigation are represented in this
figure. Annotations represent major coastal settlements along the
GBR.](../figs/log_self_strength_input_map.png){#sfig-self-conn-map fig-align="center"
width="14.82cm"}

## ADRIA-CoralBlox

The CoralBlox model incorporates linear extension (growth) and background mortality rates
that are dependent on the size and functional group of a coral colony. Environmental
disturbances and reproduction are not directly considered within CoralBlox. Instead, these
factors are considered by making changes to the distribution of colonies within CoralBlox
timesteps using the ADRIA (Adaptive Dynamic Reef Intervention Algorithms) platform
[@iwanaga2023]. Within ADRIA, we incorporate coral life history and environmental events in
the following order: coral colony growth; coral reproduction; larval connectivity and
settlement; and coral bleaching due to heat stress. As ADRIA and CoralBlox are coupled
together to represent the system, we refer to both as ADRIA-CoralBlox.

Coral fecundity rates are dependent on colony size and functional group following the
results of Hall & Hughes [-@hall1996]. Coral fertilization rates are not considered within
this model, however larval survival and viability are incorporated in connectivity modelling
(see Connectivity modelling). Viable larvae settle at a reef location and are incorporated
into the model as the smallest size class but are not directly impacted by modelled
bleaching mortality. There has been contrasting evidence for juvenile corals being immune to
thermal bleaching events [@depczynski2013], we assume the high background mortality of
juveniles includes thermal stress [@álvarez-noriega2018].

### Scenario generation

Scenarios were sampled from the range of environmental and ecological factors using the
low-discrepancy Sobol' sampling method [@sobol1967; @sobol1993] with Owen scrambling
[@owen1995]. A base of 2 and padding of 32 is used for the Owen scrambling. Ecological
factors were sampled using a +/- 10% range relative to their nominal (assumed best guess)
value.

### Model parameterisation

In this investigation we model five functional groups common to the GBR, tabular Acropora,
corymbose Acropora, corymbose non-Acropora, small massives and large massives. We do not
model arborescent Acropora forms due to their distinct life history from other morphological
groups. CoralBlox uses user-defined size classes that influence size-dependent growth and
mortality rates, through which the coral colonies progress as they grow in diameter. Within
ADRIA, we define the distributions of these parameters, and the size classes based on
ecological monitoring data from the Ecological Intelligence for Reef Restoration Program
(EcoRRAP) ([@australianinstituteofmarinescience2025; @ribeirodealmeida2024]).

In ADRIA we track the thermal tolerance of each functional group and size class as a
truncated normal distribution that is updated in each timestep following growth and
recruitment. This distribution determines the proportion of the population on a reef that is
susceptible to bleaching (i.e. has a thermal threshold below the experienced DHW value).
Starting distributions for thermal tolerance levels are informed by Bairos-Novak et al.
[-@bairos-novak2021].

### Time series clustering

The CID time series clustering metric corrects the Euclidean distance between pairs of time
series (which accounts for magnitude differences) with the complexity of each time series
(obtained through a measure of their variability) to ensure reefs are clustered based on
their underlying temporal dynamics rather than simply their average coral cover. The process
creates a distance matrix of n \* n dimensions (where n is the number of time series in the
group) that is clustered using a k-medoids clustering algorithm [cf. @steinmann2020]. The
three reef clusters are labelled with qualitative-ordered clusters from Low, Medium and High
based on their median cover relative to each other cluster over the 2030-2060 period. The
selected time frame prevents the classification from being overly influenced by any steep
declines in the initial years of simulations and avoids the post-2060 time period where
reefs are projected to be under severe stress and in a declined state [@bozec2025].

## ACCESS-ESM1-5

![Bar plot indicating the Spearman's rank correlation values for the number of years above
carbonate budget threshold and, median reef depth and log10 weighted incoming connectivity
values. Depth correlation values are displayed in the left plot, while connectivity
correlation values are displayed in the right plot. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/depth_conn_carbonate_correlations.png){#sfig-access-esm1-5-carb-budget-corr
fig-align="center" width="14.82cm"}

![Scatter plot indicating reef median depth (metres) and log10 weighted incoming
connectivity for each reef. Point colours represent the median number of years each reef
maintains positive carbonate budgets across each $theta$ value from 10-20% live coral cover.
Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/depth_conn_mean_carbonate_budget.png){#sfig-access-esm1-5-mean-carb-budget
fig-align="center" width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by median number of years exceeding positive carbonate
budget threshold. Colour-coded areas outline GBR marine park management regions, where
orange represents the Far Northern management region, green represents Cairns/Cooktown, red
covers Townsville/Whitsunday and blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids
are marked individually for visual clarity. Reefs that have a median positive carbonate
budget year value of 0 years are coloured in grey. All other reefs are coloured according to
their median years across threshold values, with darker colours representing greater number
of years with positive carbonate budgets than lighter colours. All 2197 reefs included in
this investigation are represented in this figure. This map depicts values from the
ACCESS-ESM1-5 GCM. Annotations represent major coastal settlements along the
GBR.](../figs/ACCESS-ESM1-5/median_carbonate_budget_map.png){#sfig-carb-budget-map
fig-align="center" width="14.82cm"}

### Bioregion scale clustering

![Reef area-relative coral cover time series clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2060 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2060 within each
bioregion. Solid lines represent the cluster median time series and bands represent the 95th
percentile confidence intervals. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/bioregion_cover_timeseries.png){#sfig-access-esm1-5-bioregion-timeseries
fig-align="center" width="14.82cm"}

![Mean time series DHW values for reefs within each bioregion, separated into clusters
assigned by time series clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2060 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2060
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2060 within each bioregion. Solid lines represent the
median DHW level for each cluster and bands represent the distribution of mean DHW levels
within each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwmean
fig-align="center" width="14.82cm"}

![Reef carrying capacity (coral habitable area, in base-10 log scale) for reefs within each
bioregion, separated into clusters assigned by time series clustering. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2060 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2060 within each
bioregion. Solid lines represent the median carrying capacity for each cluster and bands
represent the distribution of values within each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/carrying_capacity_bioregion_violin.png){#sfig-access-esm1-5-bioregion-carrying-capacity
fig-align="center" width="14.82cm"}

![Source-to-sink connectivity ratios (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median source-to-sink ratio for each cluster and
bands represent the distribution of source-to-sink ratios within each cluster. The vertical
grey line represent a source-to-sink ratio of 0 (in base-10 log scale) indicating equal
outgoing and incoming connectivity levels. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi
fig-align="center" width="14.82cm"}

![Weighted incoming connectivity (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median base-10 log of weighted incoming connectivity
for each cluster and bands represent the distribution of connectivity values within each
cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/weighted_incoming_conn_bioregion_violin.png){#sfig-access-esm1-5-bioregion-weightedincoming
fig-align="center" width="14.82cm"}

![Outgoing connection strength (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median base-10 log of outgoing connection strength
for each cluster and bands represent the distribution of connectivity values within each
cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/outgoing_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-outgoingconn
fig-align="center" width="14.82cm"}

![Larval retention probability (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median base-10 log of larval retention probability
for each cluster and bands represent the distribution of probability values within each
cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/self_conn_bioregion_violin.png){#sfig-access-esm1-5-bioregion-selfconn
fig-align="center" width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by mean time series DHW value. Colour-coded areas
outline GBR marine park management regions, where orange represents the Far Northern
management region, green represents Cairns/Cooktown, red covers Townsville/Whitsunday and
blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are marked individually for
visual clarity. Darker colours represent higher mean DHW values than lighter colours. All
2197 reefs included in this investigation are represented in this figure. This map depicts
values from the ACCESS-ESM1-5 GCM. Annotations represent major coastal settlements along the
GBR.](../figs/ACCESS-ESM1-5/mean_dhw_input_map.png){#sfig-mean-dhw-map fig-align="center"
width="14.82cm"}

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by weighted incoming connectivity levels (Log10).
Colour-coded areas outline GBR marine park management regions, where orange represents the
Far Northern management region, green represents Cairns/Cooktown, red covers
Townsville/Whitsunday and blue covers Mackay/Capricorn [@GBRMPA2007]. Reef centroids are
marked individually for visual clarity. Darker colours represent higher weighted incoming
connectivity values than lighter colours. All 2197 reefs included in this investigation are
represented in this figure. This map depicts values from the ACCESS-ESM1-5 GCM. Annotations
represent major coastal settlements along the
GBR.](../figs/ACCESS-ESM1-5/weighted_incoming_conn_input_map.png){#sfig-incoming-conn-map
fig-align="center" width="14.82cm"}